Probability

### 2016-04-24 prob.tex
prob.tex \section{Quantum Probability Spaces} fix a typo(?)
prob.tex "Classical Probability Spaces" Fix citation and math symbols(?)

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can be computed but with increasingly large resources as the
precision increases. Therefore it is more appropriate in such
situations to consider probability measures mapping events to a finite
set of elements computable with a fixed set of
resources~\cite{Artin1991,Wechler1992,PurdomBrown1995,DummitFoote2004}. The
simplest such set, and the one we will consider exclusively in this
paper, is the set
$\mathscr{L}_{2}=\left\{
  \textrm{impossible},\textrm{possible}\right\}$
together with the operation $\vee$ where
$x \vee y=\textrm{impossible}$ if and only if
$x=y=\textrm{impossible}$. In relation to the first definition, one
can interpret impossible as the closed interval $[0]$, possible as the
open interval $(0,1]$, and $\vee$ as the union of intervals. The
definition of a probability measure in this case is modifed as being a
function $\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \textrm{possible}$, and 
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

@@

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)= \ip{\psi}{E\psi} \label{eq:Born}
\end{equation}
which is called the Born rule~\cite{Born1984,Mermin2007}.\\
\yutsung{One small thing is whether we should say
\begin{quote}
for each normalized state~$\ket{\psi}$...$\mu_{\psi}(E)=\ip{\psi}{E\psi}$
\end{quote}
as in Mermin~\cite{Mermin2007} and Maassen~\cite{Maassen2010}
or
\begin{quote}
for each state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $...$\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}$
\end{quote}
as in Abramsky~\cite{DBLP:journals/corr/abs-0910-2393}. Personally,
I prefer Abramsky's approach, because normalizing a state in DQT is
messier than in CQT, and we have already considered states whose norm
is not one in our previous paper~\cite{DQT2014}.}\\
Our previous examples can be verified in terms of $\mu_{\psi}$. Because
$\proj{0}$ and $\proj{-}$ are non-commutative, their total probabilities
are not a probability of an event
\begin{eqnarray*}
\mu_{0}(\proj{0})+\mu_{0}(\proj{-}) & = & \frac{3}{2}>1\textrm{ .}
\end{eqnarray*}
When events are orthogonal, i.e., commutative, they can be measured
together and their probability is additive: 
\[
\mu_{0}(\proj{0})+\mu_{0}(\proj{1})=\mu_{0}(\proj{+})+\mu_{0}(\proj{-})=1=\mu_{0}(\mathbb{1})\textrm{ .}
\]
Similarly to the classical case, the total probability is one and
the orthogonal additivity can be summarized into the following two
rules:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_{i}$, we have $\mu(\sum_{i}E_{i})=\sum_{i}\mu(E_{i})$.
\end{itemize}
These two conditions define quantum probability measures. Recall that
the Born rule (\ref{eq:Born}) induces a probability measure from
a state. Naturally, we may ask: is every probability measure induced
from a state by the Born rule? The answer is yes by Gleason's theorem
when the dimension~$\ge3$~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.

>>>>

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can be computed but with increasingly large resources as the
precision increases. Therefore it is more appropriate in such
situations to consider probability measures mapping events to a finite
set of elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The
simplest such set, and the one we will consider exclusively in this
paper, is the set
$\mathscr{L}_{2}=\left\{
  \textrm{impossible},\textrm{possible}\right\}$
together with the operation $\vee$ where
$x \vee y=\textrm{impossible}$ if and only if
$x=y=\textrm{impossible}$. In relation to the first definition, one
can interpret impossible as the closed interval $[0,0]$, possible as the
open interval $(0,1]$, and $\vee$ as the addition of intervals. The
definition of a probability measure in this case is modifed as being a
function $\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \textrm{possible}$, and 
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

@@

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)= \ip{\psi}{E\psi} \label{eq:Born}
\end{equation}
which is called the Born rule~\cite{Born1984,Mermin2007}.\\
\yutsung{One small thing is whether we should say
\begin{quote}
for each normalized state~$\ket{\psi}$...$\mu_{\psi}(E)=\ip{\psi}{E\psi}$
\end{quote}
as in Mermin~\cite{Mermin2007} and Maassen~\cite{Maassen2010}
or
\begin{quote}
for each state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $...$\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}$
\end{quote}
as in Abramsky~\cite{DBLP:journals/corr/abs-0910-2393}. Personally,
I prefer Abramsky's approach, because normalizing a state in DQT is
messier than in CQT, and we have already considered states whose norm
is not one in our previous paper~\cite{DQT2014}.}\\
Our previous examples can be verified in terms of $\mu_{\psi}$. Because
$\proj{0}$ and $\proj{-}$ are non-commutative, their total probabilities
are not a probability of an event
\begin{eqnarray*}
\mu_{0}(\proj{0})+\mu_{0}(\proj{-}) & = & \frac{3}{2}>1\textrm{ .}
\end{eqnarray*}
When events are orthogonal, i.e., commutative, they can be measured
together and their probability is additive: 
\[
\mu_{0}(\proj{0})+\mu_{0}(\proj{1})=\mu_{0}(\proj{+})+\mu_{0}(\proj{-})=1=\mu_{0}(\mathbb{1})\textrm{ .}
\]
Similarly to the classical case, the total probability is one and
the orthogonal additivity can be summarized into the following two
rules:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_{i}$, we have $\mu(\sum_{i}E_{i})=\sum_{i}\mu(E_{i})$.
\end{itemize}
These two conditions define quantum probability measures. Recall that
the Born rule (\ref{eq:Born}) induces a probability measure from
a state. Naturally, we may ask: is every probability measure induced
from a state by the Born rule? The answer is yes by Gleason's theorem
when the dimension~$\ge3$ and extend the domain of the Born rule
to mixed state~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
